---
title: "Climate Change Data Exploration Tidyverse"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

# Obtaining the Data
We have to specify the column types, as by default readr will believe 'Per Capita' to be an integer when infact it needs to be a double. 

```{r}
climate<-read_csv("https://pkgstore.datahub.io/core/co2-fossil-by-nation/fossil-fuel-co2-emissions-by-nation_csv/data/0f04181960a0a896ebaf6d8afb0b71a6/fossil-fuel-co2-emissions-by-nation_csv.csv",
                  col_types = cols(
                    Year = col_integer(),
                    Country = col_character(),
                    Total = col_integer(),
                    `Solid Fuel` = col_integer(),
                    `Liquid Fuel` = col_integer(),
                    `Gas Fuel` = col_integer(),
                    Cement = col_integer(),
                    `Gas Flaring` = col_integer(),
                    `Per Capita` = col_double(),
                    `Bunker fuels (Not in Total)` = col_integer())
)
```

# Inspecting the top pollutors, per year and all time.
Some basic aggregation validates the received wisdom that the US has historically been the biggest contributor to global warming, but that China is now a bigger issue.
```{r}

climate %>% top_n(20, Total) %>% arrange(desc(Total))

top10alltime <- climate %>% 
         group_by(Country) %>%
         summarise(total = sum(Total)) %>%
         arrange(desc(total)) %>% 
         top_n(10)

top10alltime

```

```{r, fig.width=12}
climate %>% 
  filter(Country %in% top10alltime$Country) %>%
  filter(Year>2000) %>%
  ggplot(aes(x=Year, y =Total)) +
    geom_line() +
    theme(legend.position="none") +
    facet_grid(. ~ Country) +
    ggtitle("Total co2 Emissions of Largest 10 Polluters" , subtitle = "2000-2014")
```

# Trends in co2 emmissions origin - big picture
```{r}
climate %>% 
  gather(key=fueltype, Emissions, `Solid Fuel`:`Gas Flaring`) %>% 
  ggplot(aes(x=Year, y=Emissions, color = fueltype)) + geom_bar(stat="identity")

```

# Visualising recent changes - total
```{r, fig.height=13, fig.width=5}
climate %>% 
    filter(Year>2000) %>%
    gather(key=fueltype, Emissions, `Solid Fuel`:`Gas Flaring`) %>% 
    group_by(fueltype, Year) %>%
    summarise(emissions = sum(Emissions)) %>%
    ggplot(aes(x=Year, y=emissions, fill = fueltype)) + geom_bar(stat="identity") + facet_grid(fueltype ~ ., scales = "free_y")

```

# Specific Changes by Emission Source of the largest pollutors.
We let the y axis scale per emission source, worth keeping an eye on the overall scale.
```{r, fig.width=10}
climate %>% 
    filter(Year>2000) %>%
    filter(Country %in% top10alltime$Country) %>%
    gather(key=fueltype, Emissions, `Solid Fuel`:`Gas Flaring`) %>% 
    ggplot(aes(x=Year, y=Emissions, fill = fueltype)) +
        geom_bar(stat="identity") + 
        facet_grid(fueltype ~ Country, scales = "free_y", labeller = label_wrap_gen(10)) +
        theme(legend.position = "none", axis.text.x = element_text(angle = 90))
```